% Section 1.1 Anatomy of Figures

\documentclass[nooutcomes]{ximera}
%\documentclass[space,handout,nooutcomes]{ximera}

\input{../preamble.tex}

\title{Measuring by Sight}
\author{Brad Findell}
\begin{document}
\begin{abstract}
Short-answer questions involving measuring. 
\end{abstract}
\maketitle

\section{Careful Measurement by Sight}
Adjust the figures to fit the given conditions within \textbf{eyeball accuracy}.  Enter the requested measurements.  

% Perpendicular
\begin{problem}
\begin{center}  
\geogebra{gjf28er6}{450}{250}  
\end{center}
In figure above, when point $C$ is adjusted so that $\overline{BC}$ is perpendicular to $\overline{AC}$, $AC=\answer[tolerance=.07]{2.09}$.
\begin{hint}
When two lines are \emph{perpendicular}, they cross to create four congruent angles. 
\end{hint}
\begin{hint}
Use the corner of a piece of paper.
\end{hint}
\end{problem}


% Lines in a triangle
\begin{problem}
\begin{center}  
\geogebra{q32gyaud}{500}{390}  
\end{center}
In $\triangle ABC$ above, move point $D$ to make the following measurements.  \textbf{Enter -1 if it is not possible.}   
\begin{enumerate}
\item When $\overline{BD}$ is a median, $AD=\answer[tolerance=.08]{2.25}$.
\begin{hint}
A median is drawn from a vertex to the midpoint of the opposite side.
\end{hint}
\item When $\overline{BD}$ is a angle bisector, $AD=\answer[tolerance=.08]{2.78}$.
\begin{hint}
An angle bisector cuts an angle in half.  Focus near the vertex of the angle rather than near $D$.
\end{hint}
\item When $\overline{BD}$ is a perpendicular bisector, $AD=\answer{-1}$.
\begin{hint}
An perpendicular bisector cuts an segment in half and is perpendicular to it. \textbf{Enter -1 if it is not possible.} 
\end{hint}
\item When $\overline{BD}$ is a altitude, $AD=\answer[tolerance=.11]{6.46}$.
\begin{hint}
An altitude contains a vertex and is perpendicular to the line containing the opposite side. \textbf{Enter -1 if it is not possible.} 
\end{hint}
\end{enumerate}
\end{problem}


% Triangle Height, v2
\begin{problem}
\begin{center}  
\geogebra{a888zyw2}{600}{220}  
\end{center}
In $\triangle ABC$ above, the height to base $\overline{AC}$ is $\answer[tolerance=.003]{3.585}$.
\begin{hint}
You may move point $D$.  A height is the length of an altitude, which must be perpendicular to the line containing the chosen base.  
\end{hint}
\end{problem}

% Triangle Height, v1
\begin{problem}
\begin{center}  
\geogebra{kta9hbuf}{500}{300}  
\end{center}
In $\triangle ABC$ above, the height to base $\overline{AC}$ is $\answer[tolerance=.003]{3.511}$.
\begin{hint}
You may move point $D$.  A height is the length of an altitude, which must be perpendicular to the line containing the chosen base.  
\end{hint}
\end{problem}





\end{document}

